A Plasma Torus Around a Young Low-Mass Star

### 2025-05-25 paper/ms.tex
update

The material's composition -- either pure plasma or dusty plasma -- can
be clarified by time-series optical and infrared spectrophotometry.
While observations of CPVs in the optical suggest chromaticity
consistent with dust \citep{Tanimoto2020,Gunther2022,Koen2023}, 
electron scattering in a plasma transiting over a
spotted background star might also produce chromatic features
\citep{Rackham2018}.  The composition and size distribution of any dust
that is present could be determined by measuring the extinction curve
for a sample of CPVs from 1-20\,$\mu$m.  Composition and size
distributions similar to debris from rocky bodies seen around white
dwarfs \citep{Reach2009} would suggest an extrinsic origin channel.
Compositions and sizes similar to the interstellar medium would suggest
that dust can condense out from M dwarf winds, similar to processes that
occur around evolved stars \citep{Marigo2008}.

The role of the star's magnetic field could be better understood through
new observations, and new theory.  From the modeling perspective,
there is a need for rigid-field magnetohydrodynamic models to
go beyond previous work \citep{Townsend2008,Krticka2022}
and to explore what field topologies explain the observed light
curve shapes.  In particular, dynamo simulations of
fully-convective M dwarfs have suggested that global-scale mean fields
might be confined to a single hemisphere \citep{Brown2020}; such fields
would yield accumulation surfaces quite different from those that have
previously been explored.  Observationally, spectropolarimetry has the
potential to assess both the field strength and topology.  An
independent constraint on magnetic field topology may also come from
radio observations, which have shown \citep{Kaur2024} that CPVs emit
variable radio signals, including persistent and polarized components.
Detecting radio emission produced by an electron cyclotron maser
instability (e.g.~\citealt{Callingham2021}) in particular would provide
a measurement of the field strength at the site of the emitting region,
which could help clarify the magnetosphere's structure.

@@

{\it \large Contributions}: 
LGB led the inception, design, execution, and interpretation of the
research.  MMJ contributed to the interpretation.

>>>>

The material's composition -- either pure plasma or dusty plasma --
can be clarified by time-series infrared spectrophotometry.  While
observations of CPVs in the optical suggest chromaticity consistent
with dust \citep{Tanimoto2020,Gunther2022,Koen2023}, electron
scattering in a plasma transiting over a spotted background star might
also produce chromatic features \citep{Rackham2018}.  \added{This
degeneracy is alleviated in the infrared, where the} composition and
size distribution of any dust that is present could be determined by
measuring the extinction curve for a sample of CPVs from 1-20\,$\mu$m.
Composition and size distributions similar to debris from rocky bodies
seen around white dwarfs \citep{Reach2009} would suggest an extrinsic
origin channel.  Compositions and sizes similar to the interstellar
medium would suggest that dust can condense out from M dwarf winds,
similar to processes that occur around evolved stars
\citep{Marigo2008}.

\added{Future models could take a range of steps to clarify the origin
of CPVs.  A first step for magnetohydrodynamic models that can
currently explain the H$\alpha$ features \citep[e.g.][]{Waugh2022}
would be to incorporate an opacity source that can also reproduce the
photometric dips.  It is likely that this effort will require
exploring what stellar magnetic field topologies reproduce the
observed light curve shapes \citep[e.g.][]{Bouma2024}. While previous
models in the rigid-field approximation have explored both dipolar
\citep{Townsend2008} and multipolar fields \citep{Krticka2022}, dynamo
simulations of fully-convective M dwarfs have suggested that
global-scale mean fields might be confined to a single hemisphere
\citep{Brown2020}; such fields would yield accumulation surfaces quite
different from those that have been explored.  One related challenge
is that CPVs currently suffer a dearth of stellar surface field
measurements, unlike benchmark systems such as AB Dor and V374~Peg
\citep{Jardine2002,Vidotto2011}. This might be alleviated through
future spectropolarimetric observations, or from radio observations
sensitive to the magnetic field structure around the star
\citep[see][]{Callingham2021,Kaur2024}.  Finally, the question of what
microphysical process yields the opaque material also merits future
theoretical attention. While recent advances can produce the
catastrophic cooling that leads to dense, cool clumps embedded in a
hot corona \citep{Daley-Yates2023,Daley-Yates2024}, a self-consistent
treatment of radiation propagating through such clumps remains to be
performed. The question of whether dust grains can nucleate in such
clumps, or whether they need to be externally sourced, deserves
careful attention.}

%The role of the star's magnetic field could be better understood through
%new observations, and new theory.  From the modeling perspective,
%there is a need for rigid-field magnetohydrodynamic models to
%go beyond previous work \citep{Townsend2008,Krticka2022}
%and to explore what field topologies explain the observed light
%curve shapes.  In particular, dynamo simulations of
%fully-convective M dwarfs have suggested that global-scale mean fields
%might be confined to a single hemisphere \citep{Brown2020}; such fields
%would yield accumulation surfaces quite different from those that have
%previously been explored.  Observationally, spectropolarimetry has the
%potential to assess both the field strength and topology.  An
%independent constraint on magnetic field topology may also come from
%radio observations, which have shown \citep{Kaur2024} that CPVs emit
%variable radio signals, including persistent and polarized components.
%Detecting radio emission produced by an electron cyclotron maser
%instability (e.g.~\citealt{Callingham2021}) in particular would provide
%a measurement of the field strength at the site of the emitting region,
%which could help clarify the magnetosphere's structure.

@@

{\it \large Contributions}: 
LGB led the inception, design, execution, and interpretation of the
project.  MMJ contributed to the interpretation.
